---
title: "Color_Score"
author: "LaurenZane"
date: "8/14/2022"
output: html_document
---

Modifed from [ES code](https://github.com/hputnam/Acclim_Dynamics/blob/master/Scripts/Photographic_Bleaching.Rmd)and [Astrangia_repo/scripts](https://raw.githubusercontent.com/JillAshey/Astrangia_repo/main/scripts/Color_Score.Rmd)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## loading required libraries 

```{r}
library("vegan")
library("ggpubr")
library("gridExtra")
library("plyr") 
library("dplyr")
library("emmeans")
library("multcompView")
library("rstatix")
library("Rmisc")
```
Reading in quality controlled version of ImageJ data.
Protocol for ImageJ Color Score Analysis found [here](https://emmastrand.github.io/EmmaStrand_Notebook/ImageJ-Analysis-Protocols/). 
```{r}
print(getwd()) # check working directory
Data <- read.csv("../Data/20220803-color-score.csv")
unique(Data$Treatment)
# unique(Data$Timepoint) 

```
Normalizing to color standards

```{r}
Data$Red.Norm.Coral <- Data$Red.Coral/Data$Red.Standard
Data$Green.Norm.Coral <- Data$Green.Coral/Data$Green.Standard
Data$Blue.Norm.Coral <- Data$Blue.Coral/Data$Blue.Standard
par(mfrow=c(1,3))
boxplot(Data$Red.Standard)
boxplot(Data$Green.Standard)
boxplot(Data$Blue.Standard)
which((Data$Red.Standard)==min((Data$Red.Standard)))
which((Data$Green.Standard)==min((Data$Green.Standard)))
which((Data$Blue.Standard)==min((Data$Blue.Standard)))
par(mfrow=c(1,3))
boxplot(Data$Red.Norm.Coral)
boxplot(Data$Green.Norm.Coral)
boxplot(Data$Blue.Norm.Coral)


```
```{r}
color.score <- as.matrix(cbind(Data$Red.Norm.Coral,Data$Green.Norm.Coral,Data$Blue.Norm.Coral)) #create matrix
rownames(color.score) <- Data$Fragment.ID #name columns in dataframe
dist <- vegdist(color.score, method="euclidean") #calculate distance matrix of color scores
PCA.color <- princomp(dist) #run principal components Analysis
summary(PCA.color) # view variance explained by PCs
Color <- as.data.frame(PCA.color$scores[,1]) #extract PC1
Color$Fragment.ID <- rownames(color.score)
Color  <- cbind(Color, Data$Treatment)
colnames(Color) <- c("Color.Score", "Fragment.ID", "Treatment")   
Color$Group <- paste(Color$Treatment)
```
Color score output as a dataframe; check for outliers in the dataset
```{r}
write.table(Color, file = "../Output/20220803_CNT_Color_Score_Output.csv", append = FALSE, quote = TRUE, sep = ",",
            eol = "\n", na = "NA", dec = ".", row.names = TRUE,
            col.names = TRUE, qmethod = c("escape", "double"),
            fileEncoding = "")
x<- Color #assign values to test data frame to look for outliers
par(mar=c(10,4,2,2)) #bottom, left, top and right margins respectively
```


```{r}
# save R base graph as a pdf to the Output folder

# open a pdf file in the Output directory
pdf(file = "../Output/20220803-CNT-color-score-PCA-boxplot")

# create a boxplot PC1 color score by treatment
boxplot(Color.Score ~ Group, data = x, lwd = 1, ylab = 'PC1Color', las=2, cex=0.8) 

# close pdf file 

dev.off()
```
Statistical Analysis and data visualization 

```{r}
# for use in plotting to avoid having to run PCA again
Color.Score <- read.csv("../Output/20220803_CNT_Color_Score_Output.csv", header=T, sep=",", na.string="NA") #read in data file
mod1 <- aov(sqrt(Color.Score+200) ~ Treatment, data=Color.Score) #run an ANOVA ## add Timepoint as factor later 
hist(residuals(mod1)) #look at normality of data
boxplot(residuals(mod1)) #look at normality of data
summary(mod1)
#posthoc analyses
marginal <- emmeans(mod1, ~ Treatment)
#letter display of posthoc results
phlett <- multcomp::cld(marginal,
          alpha=0.05,
          Letters=letters,
          adjust="tukey")
phlett <- phlett[order(phlett$Treatment),]
phlett

```
Calculating means, sample size, and standard error for timepoint and treatment groups.
```{r}
All.Means <- summarySE(Color.Score, measurevar="Color.Score", groupvars=c("Treatment"))

```

Plot data

```{r}
Color.Score_plot <- ggplot(All.Means, aes(x=Treatment, y = Color.Score)) +
  geom_line() + geom_point() +
  #facet_grid(~Species, scales = "free", labeller = species_labeller) +
  geom_errorbar(aes(ymin=Color.Score-se, ymax=Color.Score+se), width=.2) +
  xlab("Treatment") + 
  ylab("Color Score") + # y-axis label and y axis range
  theme_classic() + 
  theme(legend.position = "left") +
  geom_vline(xintercept = c(7.8), colour="black", linetype="dotted") +
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  #scale_x_discrete(breaks = c("Week1", "Week2", "Week4", "Week6", "Week8", "Week12", "Week16"), labels = c("1 week", "2 week", "4 week", "6 week", "8 week", "12 week", "16 week")) +
  theme(axis.text.x = element_text(angle = 60, vjust = 1.2, hjust = 1.3)) + #Set the text angle
  theme(panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  # Change axis line
  axis.line = element_line(colour = "black"))
Color.Score_plot
ggsave(file="../Output/20220803-CNT-Color_Score.pdf", Color.Score_plot, width = 11, height = 6, units = c("in"))




```


